---
title: "1445149 and 1430275"
author: "Marc Labe and Shane Serebro"
date: "19 August 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<h3> Introduction </h3>

Firms that experience large negative price shocks are often deemed to be bad investments in the short run owing to investor sentiment and increased volatility in the period after such price shock. However, this may not necessarily be the case as seen in Bremer and Sweeney (1991). Moreover, shares that experience positive price shocks are often considered good investments as many believe this trend will continue for the foreseeable future. In this study, an assessment is carried out to determine whether these firms experience share price reversals and therefore can be traded opportunistically and profitably. 

Furthermore, this study will add to the existing literature with regards to whether the EMH holds in the Johannesburg Stock Exchange (JSE) or does the overreaction hypothesis postulated by De Bondt and Thaler (1985) serve as a better explanation of share price movements. If the EMH does indeed hold true on the JSE, abnormal profits cannot be obtained from trading on significant price movements whereas these profits are obtainable if evidence of the overreaction hypothesis can be found on the JSE. In essence, this paper sets out to reveal whether shares that experience a sudden and significant change in price undergo a reversal in the short-term. Furthermore, it aims to determine whether investing in a company that experiences a large share price movement is a profitable investment strategy for the active trader.


<h3> Methodology </h3>

<h5> Step 1: </h5>

We import the libraries

```{r, message=FALSE, warning=FALSE}
library(googledrive)
library(readxl)
library(lubridate)
```

```{r, message=FALSE, warning=FALSE}
id <-  "1SGigXMnzubpP15Y1W18GqnlVQFiv95jN"
try(drive_download(as_id(id), overwrite = FALSE), silent = TRUE)
PriceData <- read_excel("Price-Volume-MarketCap.xlsx", 
                        sheet = "Price (D)") 
#file.remove("Price-Volume-MarketCap.xlsx")
```



```{r, message=FALSE, warning=FALSE}

NaFunction <-  function(column){
  !all(is.na(column) | column == 0)
}


colClean1 <- function(x){
  gsub(" SJ Equity", "", names(x), fixed = TRUE)
}

```



```{r}
exclP <-  sapply(PriceData, NaFunction)
PriceData <-  PriceData[exclP]
PriceData$Date <-  as.Date(PriceData$Date)
names(PriceData) <- colClean1(PriceData)
```


```{r}

lookback <- 5 
window <- 5
triggerDD <- -0.15
triggerDU <- 0.15
```

```{r}

maxDD <-  function(column, lb){
  dd <- vector(mode = "double", length = (length(column) - lb))
  for (i in (lb+1):length(column)){
    if(is.na(max(column[(i-lb):i]))){
      dd[i] <- 0
    } else{
      dd[i] <- (column[i] - max(column[(i-lb):i], na.rm = T))/max(column[(i-lb):i], na.rm = T)
    }
  }
  return(dd)
}

minDU <-  function(column, lb){
  du <- vector(mode = "double", length = (length(column) - lb))
  for (i in (lb+1):length(column)){
    if(is.na(min(column[(i-lb):i]))){
      du[i] <- 0
    } else{
      du[i] <- (column[i] - min(column[(i-lb):i], na.rm = T))/min(column[(i-lb):i], na.rm = T)
    }
  }
 return(du)
}

```


```{r}
DUdf <-  as.data.frame(sapply(PriceData[-1], minDU, lb = lookback))
DUdf <-  cbind(PriceData$Date, DUdf)
names(DUdf)[names(DUdf) == "PriceData$Date"] <- "Date"


DDdf <-  as.data.frame(sapply(PriceData[-1], maxDD, lb = lookback))
DDdf <-  cbind(PriceData$Date, DDdf)
names(DDdf)[names(DDdf) == "PriceData$Date"] <- "Date"

```


```{r}
DDdf <-  DDdf[-c(1:5), ]
DUdf <-  DUdf[-c(1:5), ]
PriceData <-  PriceData[-c(1:5), ]

```

```{r}
trigIndexDD <- lapply(DDdf[-1], function(i){
  trigDD <- vector(mode = "integer", length = 1)
  l = length(i)
  s = 0
  pos = 1
  while(s <= l){
    x = which(i[(s+1):l] <= triggerDD)[1]+s
    if(is.na(x)){
      break
    } else{
      trigDD[pos] = x
    }
    s = trigDD[pos] + window
    pos = pos + 1
  }
  return(trigDD)
})

trigIndexDU <- lapply(DUdf[-1], function(i){
  trigDU <- vector(mode = "integer", length = 1)
  l = length(i)
  s = 0
  pos = 1
  while(s <= l){
    x = which(i[(s+1):l] >= triggerDU)[1]+s
    if(is.na(x)){
      break
    } else{
      trigDU[pos] = x
    }
    s = trigDU[pos] + window
    pos = pos + 1
  }
  return(trigDU)
})

```