---
title: "1445149 and 1430275"
author: "Marc Labe and Shane Serebro"
date: "19 August 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br>
<br>
We first import the 2 packages used in the importing and cleaning of the data:


```{r, message=FALSE, warning=FALSE}
library(readxl)
library(janitor)
```

<br>
Using the "readxl" package, we import the raw data from the excel file and select the "Price (D)" sheet and all the data it contains
This is then coerced into a dataframe object.
<br>

```{r, message=FALSE, warning=FALSE}
PriceData <- read_excel("C:/Users/shane/University/2019/Research Report/Data/Price-Volume-MarketCap.xlsx", 
                        sheet = "Price (D)", 
                        na = ""
                        )

df <-  as.data.frame(PriceData)
```

<br>

The "janitor" package contains a variety of functions to clean data and remove any unnecessary variables and/or observations.
Using the 'remove_empty' function, we remove all variables (columns) that contain only "#N/A" observations.

```{r, message=FALSE, warning=FALSE}
df1 <-  remove_empty(df, "cols")
```